### 2016-07-12 Projects/DelCo_2012/Report/DelCo_2012.tex
more edits to CAFS

\begin{tabbing}
\hspace{1.5in} \= \\
CAF \> Compressed Air Foam \\
CAFS \> Compressed Air Foam System \\
ESTC \> Delaware County Emergency Services Training Center \\
MCFRS \> Montgomery County (MD) Fire and Rescue Service \\
NIST \> National Institute of Standards and Technology \\
TFT \> Task Force Tips \\
\end{tabbing}

@@

@@

To better quantify the effectiveness of the suppression agents (CAF, water), the maximum temperature differences associated with each gas cooling event for each agent from the three thermocouple trees were analyzed. Figure~\ref{fig:combined_all} compares the temperature differences at each thermocouple position (via a scatterplot) for 30 of the 88 experiments. The 30 experiments are actually pairs of experiments, 15 water and 15 CAF, which were directly comparable, see Table~\ref{tab:Gas_Cooling_Descriptions}.  These experiments included solid stream, straight streams, and narrow fog streams at both nozzle positions.  While the total number of experiments was 88, there where a number of experiments that were performed which did not have a directly comparable experiments between water and CAF. For example a set of 12 Class A solutions experiments were conducted. Gas cooling experiments were conducted with water applied on a wide fog nozzle setting, no significant differences were noted so a comparable series with wide fog CAF was not conducted. Similarly larger water flows were examined but 120 gpm was the maximum water flow that could be generated with the CAFS, so those experiments are not included in this comparison. The Mitchell report includes the data for all 88 experiments \cite{Mitchell:1}

@@

\begin{table}[!ht]
\centering
\caption{Single Story Fire Suppression Results}\label{tab:Test_Results}
\begin{tabular}{lccc}
\toprule[1.5pt]
                 & \multicolumn{2}{c}{Temperature Change}                                    & Heat Flux Change \\
Test 	         & 0.3~m Below Ceiling                 & 0.3~m  Above Floor	                 & 0.15~m Above Floor \\
\midrule
CAF~1 Hallway    & 770~$^{\circ}$C - 190~$^{\circ}$C   & 680~$^{\circ}$C - 170~$^{\circ}$C   & 70~kW/m$^2$ - 20~kW/m$^2$  \\
CAF~1 Room       & 250~$^{\circ}$C - 60~$^{\circ}$C    & 90~$^{\circ}$C - 70~$^{\circ}$C     & 25~kW/m$^2$ - 0~kW/m$^2$  \\ [.25cm]
Water~1 Hallway  & 770~$^{\circ}$C - 300~$^{\circ}$C   & 760~$^{\circ}$C - 180~$^{\circ}$C   & 40~kW/m$^2$ - 15~kW/m$^2$  \\
Water~1 Room     & 180~$^{\circ}$C - 80~$^{\circ}$C    & 80~$^{\circ}$C - 70~$^{\circ}$C     & 25~kW/m$^2$ - 0~kW/m$^2$  \\ [.25cm]
CAF~2 Hallway    & 690~$^{\circ}$C - 290~$^{\circ}$C   & 690~$^{\circ}$C - 110~$^{\circ}$C   & 65~kW/m$^2$ - 10~kW/m$^2$  \\
CAF~2 Room       & 220~$^{\circ}$C - 80~$^{\circ}$C    & 110~$^{\circ}$C - 70~$^{\circ}$C    & 30~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~2 Hallway  & 770~$^{\circ}$C - 190~$^{\circ}$C   & 680~$^{\circ}$C - 170~$^{\circ}$C   & 70~kW/m$^2$ - 20~kW/m$^2$  \\
Water~2 Room     & 250~$^{\circ}$C - 50~$^{\circ}$C    & 80~$^{\circ}$C - 70~$^{\circ}$C     & 20~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~3 Hallway  & 770~$^{\circ}$C - 190~$^{\circ}$C   & 680~$^{\circ}$C - 170~$^{\circ}$C   & 60~kW/m$^2$ - 10~kW/m$^2$  \\
Water~3 Room     & 260~$^{\circ}$C - 180~$^{\circ}$C   & 90~$^{\circ}$C - 20~$^{\circ}$C     & 30~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
CAF~3 Hallway    & 880~$^{\circ}$C - 300~$^{\circ}$C   & 840~$^{\circ}$C - 230~$^{\circ}$C   & 80~kW/m$^2$ - 10~kW/m$^2$  \\
CAF~3 Room       & 190~$^{\circ}$C - 60~$^{\circ}$C    & 80~$^{\circ}$C - 60~$^{\circ}$C     & 15~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~4 Hallway  & 860~$^{\circ}$C - 430~$^{\circ}$C   & 870~$^{\circ}$C - 210~$^{\circ}$C   & 125~kW/m$^2$ - 10~kW/m$^2$  \\
Water~4 Room     & 260~$^{\circ}$C - 160~$^{\circ}$C   & 120~$^{\circ}$C - 50~$^{\circ}$C    & 15~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
CAF~4 Hallway    & 620~$^{\circ}$C - 270~$^{\circ}$C   & 550~$^{\circ}$C - 220~$^{\circ}$C   & 50~kW/m$^2$ - 15~kW/m$^2$  \\
CAF~4 Room       & 190~$^{\circ}$C - 120~$^{\circ}$C   & 120~$^{\circ}$C - 180~$^{\circ}$C   & 20~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~5 Hallway  & 820~$^{\circ}$C - 380~$^{\circ}$C   & 420~$^{\circ}$C - 90~$^{\circ}$C    & 65~kW/m$^2$ - 20~kW/m$^2$  \\
Water~5 Room     & 400~$^{\circ}$C - 140~$^{\circ}$C   & 120~$^{\circ}$C - 90~$^{\circ}$C    & 20~kW/m$^2$ - 15~kW/m$^2$  \\ [.25cm]
CAF~5 Hallway    & 650~$^{\circ}$C - 420~$^{\circ}$C   & 690~$^{\circ}$C - 20~$^{\circ}$C    & 40~kW/m$^2$ - 25~kW/m$^2$  \\
CAF~5 Room       & 390~$^{\circ}$C - 210~$^{\circ}$C   &                                     & 10~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~6 Hallway  & 800~$^{\circ}$C - 150~$^{\circ}$C   & 650~$^{\circ}$C - 20~$^{\circ}$C    & 60~kW/m$^2$ - 10~kW/m$^2$  \\
Water~6 Room     & 200~$^{\circ}$C - 90~$^{\circ}$C    & 60~$^{\circ}$C - 40~$^{\circ}$C     & 5~kW/m$^2$ - 0~kW/m$^2$  \\ [.25cm]
CAF~6 Hallway    & 760~$^{\circ}$C - 180~$^{\circ}$C   & 340~$^{\circ}$C - 30~$^{\circ}$C    & 50~kW/m$^2$ - 10~kW/m$^2$  \\
CAF~6 Room       & 180~$^{\circ}$C - 130~$^{\circ}$C   &                                     & 5~kW/m$^2$ - 0~kW/m$^2$  \\ [.25cm]
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

>>>>

\begin{tabbing}
\hspace{1.5in} \= \\
CAF \> Compressed Air Foam \\
CAFS \> Compressed Air Foam System \\
Cal Poly \> California Polytechnic State University, San Luis Obispo \\
ESTC \> Delaware County Emergency Services Training Center \\
MCFRS \> Montgomery County (MD) Fire and Rescue Service \\
NIST \> National Institute of Standards and Technology \\
TFT \> Task Force Tips \\
\end{tabbing}

@@

This study, conducted by NIST, was part of a larger research project led by California Polytechnic State University (Cal Poly), San Luis Obispo in collaboration with the Montgomery County (MD) Fire and Rescue Service.  This study was supported in part by a DHS/FEMA Assistance to Firefighters Research and Development Grant.

@@

To better quantify the effectiveness of the suppression agents (CAF, water), the maximum temperature differences associated with each gas cooling event for each agent from the three thermocouple trees were analyzed. Figure~\ref{fig:combined_all} compares the temperature differences at each thermocouple position (via a scatterplot) for 30 of the 88 experiments. The 30 experiments are actually pairs of experiments, 15 water and 15 CAF, which were directly comparable, see Table~\ref{tab:Gas_Cooling_Descriptions}. These experiments included solid stream, straight streams, and narrow fog streams at both nozzle positions. While the total number of experiments was 88, there where a number of experiments that were performed which did not have a directly comparable experiment between water and CAF. For example a set of 12 Class A solution experiments were conducted. Gas cooling experiments were conducted with water applied on a wide fog nozzle setting, and since no significant differences were noted compared to narrow fog a comparable series with wide fog CAF was not conducted. Similarly, larger water flows were examined but 120 gpm was the maximum water flow that could be generated with the CAFS, so those experiments are also not included in this comparison. The Mitchell report includes the data for all 88 experiments \cite{Mitchell:1}.

@@

\begin{table}[!ht]
\centering
\caption{Single Story Fire Suppression Results}\label{tab:Test_Results}
\begin{tabular}{lccc}
\toprule[1.5pt]
                 & \multicolumn{2}{c}{Temperature Change}                                    & Heat Flux Change \\
Test \# \& Nozzle	         & 0.3~m Below Ceiling                 & 0.3~m  Above Floor	                 & 0.15~m Above Floor \\
\midrule
CAF~1 Hallway    & 690~$^{\circ}$C - 180~$^{\circ}$C   & 590~$^{\circ}$C - 150~$^{\circ}$C   & 70~kW/m$^2$ - 20~kW/m$^2$  \\
CAF~1 Room       & 170~$^{\circ}$C - 80~$^{\circ}$C    & 80~$^{\circ}$C - 60~$^{\circ}$C     & 25~kW/m$^2$ - 0~kW/m$^2$  \\ [.25cm]
Water~1 Hallway  & 720~$^{\circ}$C - 220~$^{\circ}$C   & 700~$^{\circ}$C - 120~$^{\circ}$C   & 40~kW/m$^2$ - 15~kW/m$^2$  \\
Water~1 Room     & 120~$^{\circ}$C - 80~$^{\circ}$C    & 80~$^{\circ}$C - 70~$^{\circ}$C     & 25~kW/m$^2$ - 0~kW/m$^2$  \\ [.25cm]
CAF~2 Hallway    & 590~$^{\circ}$C - 290~$^{\circ}$C   & 500~$^{\circ}$C - 110~$^{\circ}$C   & 65~kW/m$^2$ - 10~kW/m$^2$  \\
CAF~2 Room       & 80~$^{\circ}$C -  70~$^{\circ}$C    & 70~$^{\circ}$C - 50~$^{\circ}$C     & 30~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~2 Hallway  & 640~$^{\circ}$C - 410~$^{\circ}$C   & 500~$^{\circ}$C - 230~$^{\circ}$C   & 70~kW/m$^2$ - 20~kW/m$^2$  \\
Water~2 Room     & 460~$^{\circ}$C - 300~$^{\circ}$C   & 200~$^{\circ}$C - 140~$^{\circ}$C   & 20~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~3 Hallway  & 610~$^{\circ}$C - 300~$^{\circ}$C   & 580~$^{\circ}$C - 170~$^{\circ}$C   & 60~kW/m$^2$ - 10~kW/m$^2$  \\
Water~3 Room     & 210~$^{\circ}$C - 130~$^{\circ}$C   & 60~$^{\circ}$C - 40~$^{\circ}$C     & 30~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
CAF~3 Hallway    & 880~$^{\circ}$C - 300~$^{\circ}$C   & 780~$^{\circ}$C - 230~$^{\circ}$C   & 80~kW/m$^2$ - 10~kW/m$^2$  \\
CAF~3 Room       & 190~$^{\circ}$C - 60~$^{\circ}$C    & 80~$^{\circ}$C - 60~$^{\circ}$C     & 15~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~4 Hallway  & 850~$^{\circ}$C - 360~$^{\circ}$C   & 850~$^{\circ}$C - 180~$^{\circ}$C   & 125~kW/m$^2$ - 10~kW/m$^2$  \\
Water~4 Room     & 190~$^{\circ}$C - 100~$^{\circ}$C   & 130~$^{\circ}$C - 70~$^{\circ}$C    & 15~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
CAF~4 Hallway    & 620~$^{\circ}$C - 240~$^{\circ}$C   & 530~$^{\circ}$C - 180~$^{\circ}$C   & 50~kW/m$^2$ - 15~kW/m$^2$  \\
CAF~4 Room       & 190~$^{\circ}$C - 120~$^{\circ}$C   & 120~$^{\circ}$C - 180~$^{\circ}$C   & 20~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~5 Hallway  & 700~$^{\circ}$C - 390~$^{\circ}$C   & 400~$^{\circ}$C - 90~$^{\circ}$C    & 65~kW/m$^2$ - 20~kW/m$^2$  \\
Water~5 Room     & 210~$^{\circ}$C - 140~$^{\circ}$C   & 120~$^{\circ}$C - 100~$^{\circ}$C   & 20~kW/m$^2$ - 15~kW/m$^2$  \\ [.25cm]
CAF~5 Hallway    & 550~$^{\circ}$C - 400~$^{\circ}$C   & 360~$^{\circ}$C - 30~$^{\circ}$C    & 40~kW/m$^2$ - 25~kW/m$^2$  \\
CAF~5 Room       & 390~$^{\circ}$C - 210~$^{\circ}$C   &                                     & 10~kW/m$^2$ - 5~kW/m$^2$  \\ [.25cm]
Water~6 Hallway  & 770~$^{\circ}$C - 160~$^{\circ}$C   & 480~$^{\circ}$C - 50~$^{\circ}$C    & 60~kW/m$^2$ - 10~kW/m$^2$  \\
Water~6 Room     & 150~$^{\circ}$C - 90~$^{\circ}$C    & 70~$^{\circ}$C - 50~$^{\circ}$C     & 5~kW/m$^2$ - 0~kW/m$^2$  \\ [.25cm]
CAF~6 Hallway    & 650~$^{\circ}$C - 190~$^{\circ}$C   & 570~$^{\circ}$C - 60~$^{\circ}$C    & 50~kW/m$^2$ - 10~kW/m$^2$  \\
CAF~6 Room       & 180~$^{\circ}$C - 130~$^{\circ}$C   &                                     & 5~kW/m$^2$ - 0~kW/m$^2$  \\ [.25cm]
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

Rekindling and the impact of the second nozzle can been seen in the time-temperature data of thermocouples within the fire room. Figure~\ref{fig:fs_instruments} shows the positions of thermocouple arrays TC A1 and TC A2. Note that TC A1 is closer to fuel and ignition.

\begin{figure}[!ht]
	\includegraphics[width=.85\columnwidth]{../Figures/Floor_Plans/PDFs/East_Structure/DelCo_2012_East_Structure_Instrumentation2}
	\caption{Schematic Indicating Thermcouple Arrays A1 and A2 for Fire Suppression Experiments}
	\label{fig:fs_instruments}
\end{figure}

For the first comparison tests in the series, CAF~1 and Water~1, the thermocouple data from TC A2 is shown in Figs.~\ref{fig:caf1_tca2} and \ref{fig:water1_tca2}, respectively. From the figures, the impact of the first suppression event (hallway nozzle on) is evident from the sharp decline in thermocouple temperatures.

\begin{figure}[!ht]
	\includegraphics[width=.85\columnwidth]{../Figures/Script_Figures/FSE_Test_1_092812_TC_A2}
	\caption{Time Series of Thermcouple Data from Array TC A2 from Single Story CAF~1 Fire Suppression Test}
	\label{fig:caf1_tca2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.85\columnwidth]{../Figures/Script_Figures/FSW_Test_2_092812_TC_A2}
	\caption{Time Series of Thermcouple Data from Array TC A2 from Single Story Water~1 Fire Suppression Test}
	\label{fig:water1_tca2}
\end{figure}

Similar to the gas cooling tests, the impact of water and CAF during fire suppression can be analyzed by comparing the temperature differences during the suppression interval at the two the thermocouple arrays in the fire room.
